---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# qzwslrm

<!-- badges: start -->
<!-- badges: end -->

The goal of qzwslrm is to provide a set of tools that are useful for doing validations of predicted breeding values using the LR-Method.

## Installation

You can install the development version of qzwslrm from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("fbzwsqualitasag/qzwslrm")
```

## Example

The following example shows how a validation of a set of predicted breeding values can be run.

```{r example, eval=FALSE}
library(qzwslrm)
## validation
validate_ebv(pvec_ebv_partial = vec_ebv_partial, pvec_ebv_whole = vec_ebv_whole)
```

The above command shows the basic use case. There are more functions that can handle different use cases. 

