---
title: "How To Use qzwslrm"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How To Use qzwslrm}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(qzwslrm)
```

## Installation
The package `qzwslrm` can be installed from Github

```r
# if (!is.element("devtools", installed.packages()) install.packages("devtools")
devtools::install_github("fbzwsqualitasag/qzwslrm")
```

## Disclaimer
This is a short introduction on how to use the package `qzwslrm`.

```{r vec-prepare, echo=FALSE, message=FALSE, warning=FALSE}
# file paths
s_ebv_path_whole <- qzwslrm_example_solani("whole")
s_ebv_path_partial <- qzwslrm_example_solani("partial")
# read data
tbl_solani_whole <- readr_solani(ps_path = s_ebv_path_whole)
tbl_solani_partial <- readr_solani(ps_path = s_ebv_path_partial)
# extract vectors
vec_ebv_whole <- tbl_solani_whole$ebv
vec_ebv_partial <- tbl_solani_partial$ebv
```


## Abstract
The package `qzwslrm` implements the computation of the EBV validation statistics using the LR-method. This requires two vectors with EBV for the same group of animals. One vector contains EBV for the group of animals estimated using the full dataset ('whole') and the second vector contains EBV for the same group of animals estimated using only a partial dataset ('partial'). 

Given two vectors `vec_ebv_whole` and `vec_ebv_partial` with EBV for the same group of animals from whole data and partial data, respectively, the following command computes a first set of validation statistics.

```{r}
l_val_result <- val_ebv_lrm(pvec_ebv_partial = tbl_solani_partial$ebv, 
                            pvec_ebv_whole = tbl_solani_whole$ebv)
```


## Usage
As shown above, the function `val_ebv_lrm()` is the central function that computes the validation statistics. The function `val_ebv_lrm()` returns a list with all validation statistics. The results can be shown using the summary function `summary_lrm()` or by converting them into a tibble.


## Results
The function `summary_lrm()` can be used to show a summary of the validation statistics

```{r}
summary_lrm(l_val_result)
```

If the results should be displayed as a table this can be done by converting it to a tibble and using then the function `knitr::kable()`. 

```{r}
tbl_lrm <- tibble_lrm(l_val_result)
knitr::kable(tbl_lrm)
```


## Scatterplot
The comparison of the two vectors of EBV can also be illustrated by a scatterplot. Such a plot can be generated using the function `scatterplot_lrm()`. 

```{r, message=FALSE, warning=FALSE}
tbl_ebv_whole <- readr_ebv(ps_path = qzwslrm_example_solani("whole"), ps_format = "table",
                            pn_ebv_col_idx = 4)
tbl_ebv_partial <- readr_ebv(ps_path = qzwslrm_example_solani("partial"), ps_format = "table",
                              pn_ebv_col_idx = 4)
p <- scatterplot_lrm(tbl_ebv_whole, tbl_ebv_partial)
print(p)
```

The above plot shows for each animal the pair of EBV from the whole and from the partial dataset. The blue line corresponds to the linear smoother which is drawn based on the points. The red line corresponds to the line with a slope equal to one which is the expected regression line for the 'whole' on the 'partial' EBV.

The function `scatterplot_lrm()` requires as input two dataframes with two columns. The first column contains the animal ID and the second column contains the EBV.
