---
title: "Development of Package qzwslrm"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Development of Package qzwslrm}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, eval=FALSE}
library(qzwslrm)
```

# Background
As described in [Legarra and Reverter (2018)](https://link.springer.com/article/10.1186/s12711-018-0426-6), validation of predicted breeding values is becoming more and more important. Their work uses a cross-validation based approach using results of successive genetic evaluations. These results consist of $\hat{u}_p$ and $\hat{u}_w$ which are vectors of the same length and contain predicted breeding values based on "partial" ($p$) and "whole" ($w$) data, respectively. 


# Cross-Validation Statistics
The following statistics are computed based on the input $\hat{u}_p$ and $\hat{u}_w$

* bias: $\overline{\hat{u}_p} - \overline{\hat{u}_w}$
* regression of whole on partial: $b_{w,p} = \frac{cov(\hat{u}_w,\hat{u}_p)}{var(\hat{u}_p)}$
* correlation between whole and partial: $r_{w,p} = \frac{cov(\hat{u}_w,\hat{u}_p)}{\sqrt{var(\hat{u}_p)var(\hat{u}_w)}}$
* regression of partial on whole: $b_{p,w} = \frac{cov(\hat{u}_w,\hat{u}_p)}{var(\hat{u}_w)}$


