---
title: "Development of Package qzwslrm"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Development of Package qzwslrm}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, eval=TRUE}
library(qzwslrm)
```

# Background
As described in [Legarra and Reverter (2018)](https://link.springer.com/article/10.1186/s12711-018-0426-6), validation of predicted breeding values is becoming more and more important. Their work uses a cross-validation based approach using results of successive genetic evaluations. These results consist of $\hat{u}_p$ and $\hat{u}_w$ which are vectors of the same length and contain predicted breeding values based on "partial" ($p$) and "whole" ($w$) data, respectively. 


# Cross-Validation Statistics
The following statistics are computed based on the input $\hat{u}_p$ and $\hat{u}_w$

* bias: $\overline{\hat{u}_p} - \overline{\hat{u}_w}$
* regression of whole on partial: $b_{w,p} = \frac{cov(\hat{u}_w,\hat{u}_p)}{var(\hat{u}_p)}$
* correlation between whole and partial: $r_{w,p} = \frac{cov(\hat{u}_w,\hat{u}_p)}{\sqrt{var(\hat{u}_p)var(\hat{u}_w)}}$
* regression of partial on whole: $b_{p,w} = \frac{cov(\hat{u}_w,\hat{u}_p)}{var(\hat{u}_w)}$


# Example Data
As shown in https://fbzwsqualitasag.github.io/qzwslrm/articles/qzwslrm_ebv_mix99.html, example data are prepared to run the most basic validation statistics. The example data consists of two results files from MiX99 containing predicted breeding values. The first result file contains EBV predicted based on a complete dataset. 

```{r}
(s_ebv_path_whole <- qzwslrm_example_solani("whole"))
```

The second result file contains EBV predicted for the same animals but based on a partial dataset. For our example, the phenotypic observations of all animals from the latest generation were set to missing (NA).

```{r}
(s_ebv_path_partial <- qzwslrm_example_solani("partial"))
```

These files can be read using the function `read_solani()`

```{r}
(tbl_solani_whole <- readr_solani(ps_path = s_ebv_path_whole))
```

The same can be done with the partial data 

```{r}
(tbl_solani_partial <- readr_solani(ps_path = s_ebv_path_partial))
```

# Validation
For the validation only the EBV vectors from the full and the partial data are required. Hence the validation function only needs the two vectors as input.

```{r}
(l_val_result <- val_ebv_lrm(pvec_ebv_partial = tbl_solani_partial$ebv, 
                             pvec_ebv_whole = tbl_solani_whole$ebv))
```


# Other Input Formats
For reasons of flexibility, EBV can be read from files having a number of different formats such as 'csv', 'csv2', 'delim' or 'table'. For testing purposes different input files are produced and stored inside of this package

```{r, eval=FALSE}
# csv
s_pkg_input_dir <- system.file("extdata", "ebvinput", package = "qzwslrm")
s_csv_input_path <- file.path(s_pkg_input_dir, "ebv_input.csv")
readr::write_csv(tbl_solani_partial, file = s_csv_input_path)

# csv2
s_csv2_input_path <- file.path(s_pkg_input_dir, "ebv_input.csv2")
readr::write_csv(tbl_solani_partial, file = s_csv2_input_path)

# delim - .txt
s_delim_input_path <- file.path(s_pkg_input_dir, "ebv_input.txt")
readr::write_delim(tbl_solani_partial, file = s_delim_input_path, delim = " ")
```

The input files included in the package can be obtained and read via

```{r}
sexdf_csv <- qzwslrm_example_input("csv")
tbl_ebv_csv <- readr_ebv(ps_path = sexdf_csv, ps_format = "csv", 
                     ps_animal_col_name = "animal", ps_ebv_col_name = "ebv")
tbl_ebv_csv
```

Testing the input format "table"

```{r}
sexdf_delim <- qzwslrm_example_input(ps_data_format = "delim")
tbl_ebv_table <- readr_ebv(ps_path = sexdf_delim, ps_format = "table",
                           ps_animal_col_name = "animal", ps_ebv_col_name = "ebv")
tbl_ebv_table
```

The generic function `readr_ebv()` for reading ebv from input files can read data from files with different formats. Furthermore, the input files can contain more information than just EBV. But the result of the reader function is always a tibble with two columns. The first column contains the animal IDs and the second the EBV of the animal.


# Summary Output
The output of the validation function is formatted using a summary function `summary_lrm()`. 

```{r}
tbl_ebv_whole <- readr_ebv(ps_path = qzwslrm_example_solani("whole"), ps_format = "table",
                            pn_ebv_col_idx = 4)
tbl_ebv_partial <- readr_ebv(ps_path = qzwslrm_example_solani("partial"), ps_format = "table",
                            pn_ebv_col_idx = 4)
summary_lrm(l_val_result <- val_ebv_lrm(pvec_ebv_partial = tbl_ebv_partial$ebv, 
                             pvec_ebv_whole = tbl_ebv_whole$ebv))
```


# Result Tibble
The results can also be converted into a tibble. This can be useful for an output as a table

```{r}
 tbl_ebv_whole <- readr_ebv(ps_path = qzwslrm_example_solani("whole"), ps_format = "table",
                            pn_ebv_col_idx = 4)
 tbl_ebv_partial <- readr_ebv(ps_path = qzwslrm_example_solani("partial"), ps_format = "table",
                              pn_ebv_col_idx = 4)
 tibble_lrm(l_val_result <- val_ebv_lrm(pvec_ebv_partial = tbl_ebv_partial$ebv,
                                         pvec_ebv_whole = tbl_ebv_whole$ebv))
```


# Scatterplots of EBV from Whole and Partial Data
The function `scatterplot_lrm()` creates a scatterplot of the EBV from whole and partial data

```{r, message=FALSE, warning=FALSE}
tbl_ebv_whole <- readr_ebv(ps_path = qzwslrm_example_solani("whole"), ps_format = "table",
                            pn_ebv_col_idx = 4)
tbl_ebv_partial <- readr_ebv(ps_path = qzwslrm_example_solani("partial"), ps_format = "table",
                              pn_ebv_col_idx = 4)
p <- scatterplot_lrm(tbl_ebv_whole, tbl_ebv_partial)
print(p)
```

The above plot shows a scatterplot of the ebv from whole data and partial data. The blue line is the linear smoothed regression line and the red line is the expected regression line with a slope of $1$. 

